\thispagestyle{empty}
\section*{Acknowledgements}
\vspace{0.5cm}
I am deeply grateful for the opportunity to conduct my research in a Japanese laboratory as part of a double-degree program. This experience has been profoundly enriching both personally and professionally. I would like to express my heartfelt thanks to Professor Yasutaka Fujimoto for welcoming me into his laboratory and providing me with the chance to meet and work with inspirational individuals, including Flavis, Besong, Brice, and Martin. Our spontaneous conversations about research and our futures were both enjoyable and enlightening.

I extend my sincere gratitude to Shun Inowe for his initial help in navigating a culture so different from my own. His assistance made my transition much smoother. I am also thankful to Eriko Hino for our daily coffee talks, which were a wonderful way to start each day with her positive humor.

I am deeply appreciative of Professor Monica Reggiani and her group, including Professor Stefano Michieletto, for their technical support during challenging times in my research.

Special thanks to Elaine, Prateek, and Hassan for their support from the company side during my thesis.

I would also like to extend my heartfelt thanks to Professor Roberto Oboe for coordinating this experience abroad and connecting me with the laboratory. His efforts made this incredible opportunity possible.

I am particularly grateful to Marco for the inspirational talks on our shared interests. Your support during difficult times helped me stay focused on my goals. I am profoundly thankful to my girlfriend, Giulia, who played a key role in this journey. Her unwavering support from day one has been indispensable.

Lastly, but not last, I thank my family, who have always been there when I needed them, and all the new friends I made in Japan. Your presence and support have made this experience unforgettable.
\afterpage{\blankpage}